---
title: 'Indicators of transparency: Figures of analysis across PMCOA'
author: "Stylianos Serghiou"
date: '`r format(Sys.time(), "%d/%m/%Y")`'
params:
  frac:
    label: "Fraction of total data to include in analysis"
    value: 0.0005
    input: slider
    min: 0
    max: 1
    step: 0.0005
output:
  prettydoc::html_pretty:
    # no code_folding available
    theme: hpstr      # or: architect; https://github.com/yixuan/prettydoc
    highlight: github # or: vignette
    toc: TRUE         # no toc_float available
    df_print: kable   # obviates %>% kable; does not replace styling though
  tufte::tufte_handout: default
  pdf_document:
    highlight: tango
    df_print: kable
    latex_engine: pdflatex
    keep_tex: yes
  rmdformats::readthedown:
    highlight: kate
    df_print: kable    # obviates %>% kable; does not replace styling though
    code_folding: hide # or: show; (comment out to not give option)
  tufte::tufte_html: 
    toc: TRUE
  epuRate::epurate:
    df_print: kable
    toc: yes
  html_document:
    highlight: tango
    theme: sandstone
    df_print: kable
    toc: yes
    toc_depth: 2
    toc_float: yes
    css: "path_to_custom.css"
header-includes:
- \DeclareUnicodeCharacter{3B8}{~}
- \DeclareUnicodeCharacter{3B1}{~}
- \DeclareUnicodeCharacter{3B2}{~}
- \DeclareUnicodeCharacter{223C}{~}
- \DeclareUnicodeCharacter{2264}{~}
- \DeclareUnicodeCharacter{2265}{~}
- \DeclareUnicodeCharacter{2581}{~}
- \DeclareUnicodeCharacter{2582}{~}
- \DeclareUnicodeCharacter{2583}{~}
- \DeclareUnicodeCharacter{2585}{~}
- \DeclareUnicodeCharacter{2586}{~}
- \DeclareUnicodeCharacter{2587}{~} 
- \DeclareUnicodeCharacter{FB00}{~} 
- \usepackage{graphicx}
editor_options: 
  chunk_output_type: inline
---

<style>
p {

text-align: justify;
text-justify: interword;
padding: 0 0 0.5em 0

}
</style>

```{r setup, include=FALSE}
# Load packages
library(knitr)
library(rmdformats)
library(kableExtra)
library(ggplot2)
library(magrittr)



######### knitr

# Define chunk options
opts_chunk$set(
  echo = T,
  cache = F,  # if TRUE, no need to rerun chunks
  # cache.lazy = TRUE,  # use with big objects (>1 GB)
  cache.comments = F,  # do not rebuild if comments change
  tidy = F,  # can play with this
  warning = F, 
  message = F,
  comment = NA,
  fig.align = "center",
  fig.width = 7,
  fig.path = "Figs/",  # export all figures to dir Figs
  linewidth = 91,
  width = 75
)


# Initiatialize hook
hook_output = knit_hooks$get("output")


# Hook to wrap output text when it exceeds 'n' using linewidth
knit_hooks$set(output = function(x, options) {
  
  if (!is.null(n <- options$linewidth)) {
    x = knitr:::split_lines(x)
    
    # wrap lines wider than 'n' 
    if (any(nchar(x) > n)) 
      x <- strwrap(x, width = n)
      x <- paste(x, collapse = "\n")
  }
  hook_output(x, options)
})


# Times a chunk and prints the time it took to run it under the chunk
# To time a chunk, include in the chunk options: {r my_chunk, timeit=TRUE}
knitr::knit_hooks$set(timeit = local({
  now = NULL
  function(before, options) {
    if (before) {
      now <<- Sys.time()
    } else {
      res = difftime(Sys.time(), now)
      now <<- NULL
      # use options$label if you want the chunk label as well
      paste('Time for this code chunk:', as.character(res))
    }
  }})
)


# For more knitr options visit: https://yihui.name/knitr/options/
# and his github page: https://github.com/yihui/knitr-examples



######### kableExtra

options(knitr.kable.NA = ''  # replace NAs in tables with blank
        , digits = 3)          # round digits (doesn't work without this here!)

## Example use
# data.frame(x = c(1,2,3), y = c(4,5,6), z = c(7,8,9)) %>% 
#   kable(booktabs = T) %>% kable_styling()

# Function to simplify table styling
sable <- function(tab, escape = T, full_width = F, drop = F, font_size = 12) {
  if (drop) {
    tab %>%
      kable(escape = escape, booktabs = T) %>%
      collapse_rows(valign = "top") %>% 
      kable_styling("striped", 
                    position = "center", 
                    full_width = full_width, 
                    font_size = font_size)
  } else {
    tab %>%
      kable(escape = escape, booktabs = T) %>%
      kable_styling("striped", 
                    position = "center", 
                    full_width = full_width,
                    font_size = font_size)
  }
}

## Guidelines
# No longer need to define options(knitr.table.format = "html"). It is now automatically done as soon as you load kableExtra
# No need to run kable() every time - done automatically as soon as you load kableExtra
# Loading kableExtra nullifies any styling applied by df_table: kable in the preamble - if you are content with standard formatting, DO NOT load kableExtra



#########  ggplot2

# Set up preferred theme in ggplot2
my_theme <- 
  # this and theme_minimal() are my favorite
  theme_light() +  
  theme(
    text = element_text(color = "grey20"),
    title = element_text(face = "bold"),
    legend.key = element_rect(colour = NA, fill = NA),  # Avoid borders
    panel.border = element_blank(),
    axis.ticks = element_blank()
  )

# Make the above theme the default theme
original_theme <- theme_set(my_theme)

# Use ggsave to save plots after plotting - this reduces size dramatically



######### Live preview

# Preview the HTML without having to re-knit
# https://www.r-bloggers.com/create-r-markdown-reports-and-presentations-even-better-with-these-3-practical-tips/amp/

# xaringan::infinite_moon_reader()




######### Tabbed sections

# You can organize content using tabs by applying the .tabset class attribute to headers within a document. This will cause all sub-headers of the header with the .tabset attribute to appear within tabs rather than as standalone sections. For example:

## Quarterly Results {.tabset}

### By Product



######### Update package

# To update the package use:
# Replace ~/serghiouTemplates/inst/rmarkdown/templates/report/skeleton.rmd
# library(devtools); setwd("/Users/Stelios/"); install("serghiouTemplates")
```


# Setup {.tabset}

```{r setup_notrun}
# Load packages
library(cowplot)
library(magrittr)
library(readxl)
library(tidyverse)
library(vroom)

# Import data
meta <- vroom("../../data/tidy_data/pmcoa-meta-data_select.csv")
pmcoa <- vroom("../../data/tidy_data/indicators_all.csv")
```

```{r, include=FALSE}
if (params$frac < 1) {
  message(paste0("Using ", params$frac * 100, "% of dataset."))
  pmcoa %<>% sample_frac(size = params$frac, replace = F)
} else {
  message("Using full dataset.")
}
```


Variable groupings and definitions.

```{r}
num_vars <- syms(c(
  "score",
  "year",
  "jif",
  "eigenfactor_score",
  "n_cite"
))

num_vars_meta <- syms(c(
  "n_auth",
  "n_affiliation",
  "n_ref",
  "n_fig",
  "n_table"
))

num_vars_dic <- c(
  "Indicator count" = "score",
  "Year" = "year",
  "Journal Impact Factor (WOS)" = "jif",
  "Eigenfactor (WOS)" = "eigenfactor_score",
  "Citation count (OCC)" = "n_cite"
)

num_vars_meta_dic <- c(
  "Number of authors" = "n_auth",
  "Number of affiliations" = "n_affiliation",
  "Number of references" = "n_ref",
  "Number of figures" = "n_fig",
  "Number of tables" = "n_table"
)


factor_vars <- syms(c(
  "journal",
  "publisher",
  "affiliation_country",
  "type",
  "field",
  "is_art"
))

factor_vars_meta <- syms(c(
  "is_supplement"
))

factor_vars_dic <- c(
  "Journal" = "journal",
  "Publisher" = "publisher",
  "Country of affiliation" = "affiliation_country",
  "Type of publication" = "type",
  "Field of science (SciTech)" = "field",
  "Research article (OCC)" = "is_art"
)

factor_vars_meta_dic <- c(
  "Has supplement" = "is_supplement"
)


indicator_vars <- syms(c(
  "is_data_pred",
  "is_code_pred",
  "is_coi_pred",
  "is_fund_pred",
  "is_register_pred"
))

indicator_vars_dic <- c(
  "Data sharing" = "is_data_pred",
  "Code sharing" = "is_code_pred",
  "COI disclosure" = "is_coi_pred",
  "Funding disclosure" = "is_fund_pred",
  "Protocol registration" = "is_register_pred"
)


filter_vars <- syms(c(
  "year"
))


dic <- c(
  score = "Indicator count",
  year = "Year",
  jif = "Journal Impact Factor (WOS)",
  eigenfactor_score = "Eigenfactor (WOS)",
  n_cite = "Citation count (OCC)",
  n_auth = "Number of authors",
  n_affiliation = "Number of affiliations",
  n_ref = "Number of references",
  n_fig = "Number of figures",
  n_table = "Number of tables",
  journal = "Journal",
  publisher = "Publisher",
  affiliation_country = "Country of affiliation",
  type = "Type of publication",
  field = "Field of science (SciTech)",
  is_art = "Research article (OCC)",
  is_supplement = "Has supplement",
  is_data_pred = "Data sharing",
  is_code_pred = "Code sharing",
  is_coi_pred = "COI disclosure",
  is_fund_pred = "Funding disclosure",
  is_register_pred = "Protocol registration"
)
```


Change those that report that the study was not registered to false. The wordy syntax is required to avoid the problem of NAs.

```{r}
pmcoa %<>% 
  mutate(is_register_pred = case_when(
      not_registered_1 ~ F, 
      TRUE ~ is_register_pred
  ))
```

```{r}
pmc <- 
  pmcoa %>% 
  filter(is_success) %>% 
  select(
    pmcid_pmc, 
    !!! filter_vars, 
    !!! indicator_vars, 
    !!! num_vars, 
    !!! factor_vars
  )
```


***


# Fig 3: Indicators and by year

## Prevalence

```{r}
# Set colors
cols <- setNames(scales::hue_pal()(5), names(indicator_vars_dic))

# Table
indicator_counts_by_pmc <- 
  pmcoa %>% 
  filter(is_success) %>% 
  select(ends_with("_pred"), starts_with("is_coi"), starts_with("is_fund")) %>%
  gather(indicator, vals, !!!indicator_vars) %>% 
  mutate(is_pmc = case_when(
    indicator == "is_coi_pred" ~ is_coi_pmc_fn,
    indicator == "is_fund_pred" ~ is_fund_pmc_group | is_fund_pmc_anysource
  )) %>% 
  mutate(is_pmc = if_else(is.na(is_pmc), F, is_pmc)) %>% 
  mutate(indicator = recode(indicator, !!!dic)) %>%
  group_by(indicator) %>% 
  summarise_at(vars(vals, is_pmc), list(n = sum, p = mean)) %>%
  ungroup() %>% 
  mutate(vals_p = scales::percent(vals_p, accuracy = 0.1)) %>% 
  mutate(indicator = fct_reorder(indicator, vals_n))


# Plot
p1 <- 
  indicator_counts_by_pmc %>% 
  ggplot(aes(x = indicator, y = vals_n, fill = indicator, color = indicator)) +
  geom_col(alpha = 0.75) + 
  geom_text(aes(label = vals_p), hjust = -0.1, size = 3) +
  geom_col(aes(y = is_pmc_n)) +
  scale_y_continuous(
    name = "\nNumber of articles (in millions)",
    breaks = seq(0, 3E6, 1E6),
    labels = seq(0, 3, 1),
    expand = expansion(mult = c(0, 0.2))
  ) +
  scale_fill_manual(values = cols) +
  scale_color_manual(values = cols) +
  coord_flip() +
  xlab(NULL) +
  theme(
    legend.position = "none", 
    panel.grid.major.y = element_blank()
  )
```


## Year

```{r, fig.width=11, fig.height=7}
# Table
indicator_by_year <- 
  pmc %>% 
  select(year, !!!indicator_vars) %>% 
  filter(year >= 2000) %>% 
  gather("key", "value", -year) %>% 
  mutate(key = recode(key, !!!dic)) %>% 
  count(year, key, value) %>% 
  group_by(year, key) %>% 
  mutate(p = n / sum(n)) %>% 
  filter(value) %>% 
  ungroup() %>% 
  mutate(key = fct_relevel(key, "Data sharing"))

# Plot
p2 <- 
  indicator_by_year %>% 
  ggplot(aes(x = year, y = p, color = key)) +
  geom_line(size = 0.75) +
  scale_x_continuous(breaks = seq(2000, 2020, 5), limits = c(2000, 2020)) +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
  scale_color_discrete(name = NULL) +
  labs(y = "Proportion of articles (%)\n", x = "\nYear") +
  theme(
    panel.grid.minor = element_blank(), 
    legend.position = c(0.13, 0.87)
  )
```


## Join

```{r}
cowplot::plot_grid(
  p1, p2,
  nrow = 1, 
  labels = "auto",
  rel_widths = c(2, 3),
  scale = 0.9
)

ggsave(
  "../../output/figure_output/indicator-figs_fig-3_prevalence-by-year.jpg", 
  height = 5, 
  width = 12
)
```


***


# Fig 4: Indicators by country

Create cohort.

```{r}
pmc_country_1990 <- 
  pmc %>% 
  select(pmcid_pmc, year, affiliation_country, !!!indicator_vars) %>% 
  filter(year > 1989) %>% 
  drop_na(affiliation_country) %>% 
  tidyr::separate_rows(affiliation_country, sep = "; ") %>% 
  group_by(pmcid_pmc) %>% 
  distinct(affiliation_country, .keep_all = T) %>% 
  ungroup() %>% 
  add_count(affiliation_country) 

# Get map
world_map <- ggplot2::map_data('world') %>% as_tibble

# Update China to read Hong Kong whenever it refers to Hong Kong
world_map$region[world_map$subregion == "Hong Kong"] <- "Hong Kong"

country_n <- 
  pmc_country_1990 %>% 
  group_by(affiliation_country, n) %>% 
  summarise_at(vars(!!!indicator_vars), mean) %>% 
  right_join(world_map, by = c('affiliation_country' = 'region')) %>% 
  rename(region = affiliation_country)
```


Plot.

```{r, fig.width=12, fig.height=12}
country_n %>%
  filter(region != "Antarctica") %>% 
  gather(Indicator, vals, !!!indicator_vars) %>% 
  mutate(Indicator = recode(Indicator, !!!dic)) %>% 
  mutate(Indicator = fct_relevel(Indicator, "Data sharing")) %>% 
  ggplot(aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = vals)) + 
  facet_wrap(~ Indicator, scales = "free", nrow = 3) +
  theme(
    text = element_text(color = "grey20"),
    plot.subtitle = element_text(size = 14),
    axis.ticks = element_blank(),
    axis.text.x  = element_blank(),
    axis.text.y  = element_blank(),
    panel.grid = element_blank(),
    panel.background  = element_rect(fill = "white"),
    plot.background   = element_rect(fill = "white"),
    legend.position   = c(0.78, 0.18),
    legend.background = element_blank(),
    legend.key = element_blank()
  ) +
  labs(x = NULL, y = NULL) +
  scale_fill_viridis_c(labels = scales::percent, name = "Publications (%)\n")

# Save
ggsave(
  "../../output/figure_output/indicator-figs_fig-4_indicator-by-country.jpg", 
  height = 10, 
  width = 9
)
```


***


# Fig 5: Indicator by field

Table.

```{r}
field_table <- 
  pmc %>% 
  select(year, field, !!!indicator_vars) %>% 
  drop_na(field) %>% 
  filter(year >= 1990) %>%
  rename(!!!indicator_vars_dic) %>% 
  mutate(field = fct_infreq(field)) %>% 
  gather(indicator, value, -year, -field) %>% 
  group_by(field, indicator) %>% 
  summarise_at(vars(value), list(n = sum, p = mean)) %>% 
  mutate(p_text = sprintf("%s (%.1f%%)", n, p * 100)) %>% 
  ungroup()

```


Plot.

```{r}
colfunc <- colorRampPalette(c("red", "yellow", "yellow", "green"))

field_table %>% 
  mutate(field = fct_rev(field)) %>% 
  mutate(indicator = fct_relevel(indicator, "Data sharing")) %>% 
  ggplot(aes(x = indicator, y = field, fill = p, label = p_text)) +
  geom_tile(aes(width = 0.99, height = 0.99)) +
  scale_fill_gradientn(colours = alpha(colfunc(10), 0.7)) +
  geom_text(size = 3) +
  scale_x_discrete(position = "top", expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0)) +
  labs(x = NULL, y = NULL) +
  theme(
    axis.line = element_blank(),
    legend.position = "none",
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(size = 0.1, fill = NA),
    axis.ticks = element_blank(),
    plot.margin = unit(c(0, 0, 0, 0), "cm")
  )

ggsave(
  "../../output/figure_output/indicator-figs_fig-5_field-by-indicator.jpg",
  width = 8, 
  height = 6
)
```


***


# Fig 6: Indicatory by journal and publisher

## Journal

Table - since 1990, ≥100 publications.

```{r}
indicator_by_journal <- 
  pmc %>% 
  select(year, journal, !!!indicator_vars) %>% 
  filter(year >= 1990) %>%
  add_count(journal, name = "n_publications") %>% 
  filter(n_publications > 100) %>% 
  rename(!!!indicator_vars_dic) %>% 
  gather(indicator, value, -year, -journal, -n_publications) %>% 
  group_by(journal, indicator) %>% 
  summarise_at(vars(value), list(n = sum, p = mean))
```


Plot.

```{r}
p1 <- 
  indicator_by_journal %>% 
  mutate(indicator = fct_relevel(indicator, "Data sharing")) %>% 
  ggplot(aes(x = p, color = indicator, fill = indicator)) + 
  stat_ecdf(aes(y = 1 - ..y..), geom = "area", alpha = 0.75, pad = F) +
  scale_y_continuous(
    labels = scales::percent,
    limits = c(0, 1),
    name = "Journals (%)\n"
  ) +
  scale_x_continuous(
    labels = scales::percent, 
    name = "\nPublications with indicator (%)"
  ) +
  facet_wrap(~ indicator) +
  theme(
    legend.position = "none",
    panel.grid.minor = element_blank(),
    panel.spacing = unit(1, "lines")
  )
```


## Publisher

Table - since 1990, ≥100 publications.

```{r}
indicator_by_publisher <- 
  pmc %>% 
  select(year, publisher, !!!indicator_vars) %>% 
  filter(year >= 1990) %>%
  add_count(publisher, name = "n_publications") %>% 
  filter(n_publications > 100) %>% 
  rename(!!!indicator_vars_dic) %>% 
  gather(indicator, value, -year, -publisher, -n_publications) %>% 
  group_by(publisher, indicator) %>% 
  summarise_at(vars(value), list(n = sum, p = mean))
```


Plot.

```{r}
p2 <- 
  indicator_by_publisher %>% 
  mutate(indicator = fct_relevel(indicator, "Data sharing")) %>% 
  ggplot(aes(x = p, color = indicator, fill = indicator)) + 
  stat_ecdf(aes(y = 1 - ..y..), geom = "area", alpha = 0.75, pad = F) +
  scale_y_continuous(
    labels = scales::percent,
    limits = c(0, 1),
    name = "Publishers (%)\n"
  ) +
  scale_x_continuous(
    labels = scales::percent, 
    name = "\nPublications with indicator (%)"
  ) +
  facet_wrap(~ indicator) +
  theme(
    legend.position = "none",
    panel.grid.minor = element_blank(),
    panel.spacing = unit(1, "lines")
  )
```


## Join

Table - since 1990, ≥100 publications.

```{r}
cowplot::plot_grid(
  p1, p2,
  nrow = 2, 
  labels = "auto",
  rel_heights = c(1, 1),
  scale = 0.9
)

ggsave(
  "../../output/figure_output/indicator-figs_fig-6_by-journal-publisher.jpg", 
  width = 10,
  height = 12
)
```


***


# Documentation {.tabset}

## Session Info

```{r session_info, echo=FALSE}
print(sessionInfo(), locale = F)
```


## References

```{r refs, echo=FALSE}
(.packages()) %>% sort %>% lapply(citation) %>% lapply(c) %>% unique
```
