---
title: 'Indicators of transparency: Descriptive statistics'
author: "Stylianos Serghiou"
date: '`r format(Sys.time(), "%d/%m/%Y")`'
output:
  prettydoc::html_pretty:
    # no code_folding available
    theme: hpstr      # or: architect; https://github.com/yixuan/prettydoc
    highlight: github # or: vignette
    toc: TRUE         # no toc_float available
    df_print: kable   # obviates %>% kable; does not replace styling though
  tufte::tufte_handout: default
  pdf_document:
    highlight: tango
    df_print: kable
    latex_engine: pdflatex
    keep_tex: yes
  rmdformats::readthedown:
    highlight: kate
    df_print: kable    # obviates %>% kable; does not replace styling though
    code_folding: hide # or: show; (comment out to not give option)
  tufte::tufte_html: 
    toc: TRUE
  epuRate::epurate:
    df_print: kable
    toc: yes
  html_document:
    highlight: tango
    theme: sandstone
    df_print: kable
    toc: yes
    toc_depth: 2
    toc_float: yes
    css: "path_to_custom.css"
header-includes:
- \DeclareUnicodeCharacter{3B8}{~}
- \DeclareUnicodeCharacter{3B1}{~}
- \DeclareUnicodeCharacter{3B2}{~}
- \DeclareUnicodeCharacter{223C}{~}
- \DeclareUnicodeCharacter{2264}{~}
- \DeclareUnicodeCharacter{2265}{~}
- \DeclareUnicodeCharacter{2581}{~}
- \DeclareUnicodeCharacter{2582}{~}
- \DeclareUnicodeCharacter{2583}{~}
- \DeclareUnicodeCharacter{2585}{~}
- \DeclareUnicodeCharacter{2586}{~}
- \DeclareUnicodeCharacter{2587}{~} 
- \DeclareUnicodeCharacter{FB00}{~} 
- \usepackage{graphicx}
editor_options: 
  chunk_output_type: inline
---

<style>
p {

text-align: justify;
text-justify: interword;
padding: 0 0 0.5em 0

}
</style>

```{r knitr, echo=FALSE}
# Load packages
library(knitr)
library(rmdformats)
library(kableExtra)
library(ggplot2)
library(magrittr)



######### knitr

# Define chunk options
opts_chunk$set(echo = TRUE
               , cache     = FALSE      # if TRUE, no need to rerun chunks
               #, cache.lazy = TRUE     # use when have big objects (>1 GB)
               , cache.comments = FALSE # do not rebuild if comments change
               , tidy      = FALSE      # can play with this
               , warning   = FALSE 
               , message   = FALSE
               , comment   = NA
               , fig.align = "center"
               , fig.width = 7
               # , fig.path  = "Figs/" # export all figures to dir Figs
               , linewidth = 91)

opts_knit$set(width = 75)

# Initiatialize hook
hook_output = knit_hooks$get("output")

# Hook to wrap output text when it exceeds 'n' using linewidth
knit_hooks$set(output = function(x, options) {
  
  if (!is.null(n <- options$linewidth)) {
    x = knitr:::split_lines(x)
    
    # wrap lines wider than 'n' 
    if (any(nchar(x) > n)) 
      x <- strwrap(x, width = n)
      x <- paste(x, collapse = "\n")
  }
  hook_output(x, options)
})

# For more knitr options visit: https://yihui.name/knitr/options/
# and his github page: https://github.com/yihui/knitr-examples



######### kableExtra

options(knitr.kable.NA = ''  # replace NAs in tables with blank
        , digits = 3)          # round digits (doesn't work without this here!)

## Example use
# data.frame(x = c(1,2,3), y = c(4,5,6), z = c(7,8,9)) %>% 
#   kable(booktabs = T) %>% kable_styling()

# Function to simplify table styling
sable <- function(tab, escape = T, full_width = F, drop = F, font_size = 12) {
  if (drop) {
    tab %>%
      kable(escape = escape, booktabs = T) %>%
      collapse_rows(valign = "top") %>% 
      kable_styling("striped", 
                    position = "center", 
                    full_width = full_width, 
                    font_size = font_size)
  } else {
    tab %>%
      kable(escape = escape, booktabs = T) %>%
      kable_styling("striped", 
                    position = "center", 
                    full_width = full_width,
                    font_size = font_size)
  }
}

## Guidelines
# No longer need to define options(knitr.table.format = "html"). It is now automatically done as soon as you load kableExtra
# No need to run kable() every time - done automatically as soon as you load kableExtra
# Loading kableExtra nullifies any styling applied by df_table: kable in the preamble - if you are content with standard formatting, DO NOT load kableExtra



#########  ggplot2

# Set up preferred theme in ggplot2
my_theme <- 
  # this and theme_minimal() are my favorite
  theme_light() +  
  theme(
    axis.ticks = element_blank(),
    axis.title = element_text(face = "bold"),
    axis.text.x = element_text(margin = margin(t = 3)),
    axis.text.y  = element_text(margin = margin(r = 2)),
    legend.key = element_rect(colour = NA, fill = NA),  # Avoid borders
    panel.border = element_blank(),
    panel.spacing = unit(1, "lines"),
    text = element_text(color = "grey20"),
    title = element_text(face = "bold")
  )

# Make the above theme the default theme
original_theme <- theme_set(my_theme)

# Use ggsave to save plots after plotting - this reduces size dramatically



######### Live preview

# Preview the HTML without having to re-knit
# https://www.r-bloggers.com/create-r-markdown-reports-and-presentations-even-better-with-these-3-practical-tips/amp/

# xaringan::infinite_moon_reader()




######### Tabbed sections

# You can organize content using tabs by applying the .tabset class attribute to headers within a document. This will cause all sub-headers of the header with the .tabset attribute to appear within tabs rather than as standalone sections. For example:

## Quarterly Results {.tabset}

### By Product



######### Update package

# To update the package use:
# Replace ~/serghiouTemplates/inst/rmarkdown/templates/report/skeleton.rmd
# library(devtools); setwd("/Users/Stelios/"); install("serghiouTemplates")
```


# Setup {.tabset}

```{r setup_1}
# Load packages
library(cowplot)
library(magrittr)
library(readxl)
library(tidyverse)


# Paths
fig_output <- "../../output/figure_output"


# Import data
articles <- read_csv("../../data/tidy_data/manual-indicators.csv")
altmetric <- read_csv("../../data/external_data/altmetric-data.csv")
impact <- read_csv("../../data/external_data/impact-factor.csv")
pubmed <- read_xlsx("../../data/external_data/pubmed-info.xlsx", na = "NA")
fields <- read_xlsx("../../data/external_data/stelios_pmid_info.xlsx", 1)
cites <- read_xlsx("../../data/external_data/stelios_pmid_info.xlsx", 3)
```

```{r}
# Define variable groups
indicator_vars <- syms(c(
  "isNovel",
  "isReplication",
  "FundingType",
  "hasConflictFT",
  "hasDataNoteFT",
  "hasProtocolFT",
  "hasCode"
))

indicator_vars_binary <- syms(c(
  "Novel",
  "Replication",
  "Funding",
  "COI",
  "Data",
  "Protocol",
  "Code"
))

indicator_vars_pubmed <- syms(c(
  "isReplicationNoveltyAbstract",
  "hasFundingPubMed",
  "hasConflictPubMed",
  "hasDataNoteAbstract",
  "hasProtRegInfoAbstractPub",
  "hasCodeAbstract"
))

indicator_vars_dic <- c(
  isNovel  = "Novelty",
  isReplication = "Replication",
  isReplicationNoveltyAbstract = "Novelty or Replication (PubMed)",
  FundingType = "Funding disclosure",
  hasFundingPubMed = "Funding disclosure (PubMed)",
  hasConflictFT = "COI disclosure",
  hasConflictPubMed = "COI disclosure (PubMed)",
  hasDataNoteFT = "Data sharing",
  DataShareCategory = "Data sharing",
  hasDataNoteAbstract = "Data sharing (PubMed)",
  hasProtocolFT = "Protocol registration",
  hasProtRegInfoAbstractPub = "Protocol registration (PubMed)",
  hasCode = "Code sharing",
  Data = "Data sharing",
  Code = "Code sharing",
  COI = "COI disclosure",
  Funding = "Funding disclosure",
  Protocol = "Protocol registration",
  Novel = "Novelty"
)

indicators_simplified_dic <- c(
  isNovel  = "Novel or Replication",
  isReplication = "Novel or Replication",
  isReplicationNovelty = "Novel or Replication",
  isReplicationNoveltyAbstract = "Novel or Replication",
  FundingType = "Funding disclosure",
  hasFundingPubMed = "Funding disclosure",
  hasConflictFT = "COI disclosure",
  hasConflictPubMed = "COI disclosure",
  hasDataNoteFT = "Data sharing",
  hasDataNoteAbstract = "Data sharing",
  hasProtocolFT = "Protocol registration",
  hasProtRegInfoAbstractPub = "Protocol registration",
  hasCode = "Code sharing",
  hasCodeAbstract = "Code sharing"
)


factor_vars <- syms(c(
  "isBench",
  "isEpi",
  "isOther",
  "isRCT",
  "Year",
  "is_PMC",
  "field"
))

factor_vars_dic <- c(
  field = "Field of Science (SciTech)",
  is_PMC = "On PMC",
  isBench = "Basic sciences",
  isEpi = "Health sciences",
  isOther = "Research article",
  isRCT = "Clinical Trial"
)


num_vars <- syms(c(
  "aas_score",
  "n_cites",
  "jif",
  "eigenfactor_score",
  "n_authors",
  "n_affiliations",
  "n_mesh"
))

num_vars_dic <- c(
  aas_score = "Altmetric Attention Score", 
  n_cites = "Citation count (OCC)", 
  eigenfactor_score = "Eigenfactor (WOS)",
  jif = "Journal Impact Factor (JIF)",
  n_affiliations = "Number of affiliations",
  n_authors = "Number of authors",
  n_mesh = "Number of MeSH terms"
)


coi_dic <- c(
  `0` = "No", 
  `1` = "Yes (conflicts)", 
  `2` = "Yes (no conflicts)"
)

coi_binary_dic <- c(
  `0` = "No", 
  `1` = "Yes", 
  `2` = "Yes",
  `99` = "N/A"
)


fund_dic <- c(
  `0` = "No", 
  `1` = "No funding",
  `2` = "Public", 
  `3` = "Private", 
  `4` = "NGO", 
  `5` = "Public + Private", 
  `6` = "Public + NGO", 
  `7` = "NGO + Private", 
  `8` = "Public + Private + NGO")

fund_binary_dic <- c(
  `0` = "No", 
  `1` = "Yes", 
  `2` = "Yes", 
  `3` = "Yes", 
  `4` = "Yes", 
  `5` = "Yes", 
  `6` = "Yes", 
  `7` = "Yes", 
  `8` = "Yes"
)
```


***


# Introduction

## Inputs

* Citation data - CrossRef  - 30 May, 2019
* Altmetric data - Altmetric - 30 May, 2019
* Impact factor data - Web of Science - 2019 JIF


***


# Prepare data

```{r}
glimpse(articles)
glimpse(altmetric)
glimpse(cites)
glimpse(impact)
glimpse(pubmed)
```


## Articles

Change articles before meta, because meta is created on the basis of articles.

Change protocol to align with our own definition of protocol.

```{r}
articles %<>% 
  mutate(has_suppl_note = str_detect(hasProtocol_Notes, "Suppl")) %>% 
  mutate(hasProtocolFT = case_when(
    is.na(has_suppl_note) ~ hasProtocolFT,
    has_suppl_note ~ 0,
    !has_suppl_note ~ hasProtocolFT
  )) %>% 
  select(-has_suppl_note)
```


Convert `isResearch` into logical.

```{r}
articles %<>% 
  mutate(isResearch = isResearch == "Yes") %>% 
  mutate(isEnglish = isEnglish == "Yes")
```


Create indicators for articles.

```{r}
articles %<>% 
  mutate(
    COI = hasConflictFT != 0,
    Funding = FundingType != 0,
    Data = hasDataNoteFT != 0,
    Code = hasCode != 0,
    Protocol = hasProtocolFT != 0,
    Novel = isNovel != 0,
    Replication = isReplication != 0
  )
```



## Meta

Retain only articles from which indicators were extracted and remove country because this is misleading (this is the country of the publisher).

```{r pubmed}
pubmed %<>% 
  slice(1:520) %>% 
  select(-Country, -Year) %>% 
  rename(Journal_abbrv = Journal)
```


Merge with information provided by other services.

```{r altmetric}
meta <- 
  altmetric %>% 
  select(
    pmid, 
    type, 
    is_oa, 
    aas_score = score,
    starts_with("cited_by_"), 
    starts_with("subjects"), 
    starts_with("scopus_subjects")
  ) %>% 
  unite(col = "subjects", starts_with("subjects"), sep = "; ", na.rm = T) %>% 
  unite(col = "subjects_scopus", starts_with("sco"), sep = "; ", na.rm = T) %>%
  mutate_at(vars(subjects, subjects_scopus), na_if, "") %>% 
  mutate_at(vars(pmid), as.character) %>% 
  right_join(pubmed, by = c("pmid" = "PMID")) %>% 
  mutate(aas_score = replace_na(aas_score, 0)) %>% 
  select(Sample_ID, pmid, Title:n_citation, everything())
```

```{r cites}
cites %<>% count(pmid, name = "n_cites")

meta <- 
  cites %>%
  mutate(pmid = as.character(pmid)) %>% 
  select(pmid, n_cites) %>% 
  right_join(meta) %>% 
  select(-n_cites, n_cites)
```

```{r}
meta <- 
  fields %>% 
  rename(field = pm_field) %>% 
  mutate(pmid = as.character(pmid)) %>% 
  right_join(meta) %>% 
  select(-field, field)
```


```{r jif}
meta <- 
  impact %>% 
  select(starts_with("ISSN"), jif, eigenfactor_score) %>% 
  gather("ISSN_type", "ISSN", starts_with("ISSN")) %>% 
  distinct(jif, eigenfactor_score, ISSN) %>% 
  right_join(meta, ) %>% 
  select(-(jif:ISSN), everything(), ISSN, jif, eigenfactor_score)
```

Ten of the journals with no JIF match were manually inspected to confirm that no journal with a similar title existed in Web of Science (in addition to previous manual searches in compiling the initial database). All 10 did not exist.

```{r articles}
meta <- 
  articles %>% 
  mutate_at(vars(Sample_ID), as.numeric) %>% 
  mutate(hasCodeAbstract = "No") %>% 
  select(
    Sample_ID, 
    PMCID, 
    Year:isCostEffect, 
    isEnglish, 
    isResearch,
    !!! indicator_vars,
    !!!indicator_vars_pubmed
  ) %>% 
  right_join(meta, by = "Sample_ID")
```


Create indicators.

```{r}
meta %<>% 
  mutate(
    COI = hasConflictFT != 0,
    Funding = FundingType != 0,
    Data = hasDataNoteFT != 0,
    Code = hasCode != 0,
    Protocol = hasProtocolFT != 0,
    Novel = isNovel != 0,
    Replication = isReplication != 0
  )
```


Create features.

```{r}
meta %<>% 
  mutate(
    is_PMC = !is.na(PMCID),
    n_authors = str_count(Author, ","),
    n_affiliations = str_count(Affiliation, ";"),
    n_mesh = str_count(MeSH, ";"),
    open_score = Funding + COI + Data + Protocol + Code
  )
```


***


# Fig 1: Indicators


## Indicator prevalence

Any articles (COI/Funding) and research articles (rest).

```{r}
# Plot
articles %>% 
  filter(Reviewer != "JW" & isEnglish) %>% 
  select(isResearch, !!! indicator_vars_binary) %>%
  mutate_at(vars(-COI, -Funding, -isResearch), ~ if_else(isResearch, ., NA)) %>%
  gather(key, value, -isResearch) %>% 
  drop_na(value) %>% 
  mutate(key = fct_relevel(key, "Data")) %>% 
  mutate(key = fct_relevel(key, "Protocol", after = 4)) %>% 
  mutate(key = recode(key, !!! indicator_vars_dic)) %>% 
  mutate(value = if_else(value, "Yes", "No")) %>% 
  ggplot(aes(x = value, fill = key, color = key)) +
  geom_bar(alpha = 0.75) +
  geom_text(
    aes(
      y = stat(count), 
      label = scales::percent(stat(prop), accuracy = 0.1), 
      group = key
    ), 
    stat = "count", 
    vjust = -0.5,
    size = 3
  ) +
  facet_wrap(~ key,) +
  scale_y_continuous(limits = c(0,400)) +
  theme(
    legend.position = "none",
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank()
  ) +
  labs(x = NULL, y = "Number of articles\n")
```


Research articles.

```{r}
# Plot
p1 <- 
  articles %>% 
  filter(Reviewer != "JW" & isEnglish & isResearch) %>% 
  select(!!! indicator_vars_binary) %>%
  gather(key, value) %>% 
  drop_na(value) %>% 
  mutate(key = fct_relevel(key, "Data")) %>% 
  mutate(key = fct_relevel(key, "Protocol", after = 4)) %>% 
  mutate(key = recode(key, !!! indicator_vars_dic)) %>% 
  ggplot(aes(x = value, fill = key, color = key)) +
  geom_bar(alpha = 0.75) +
  geom_text(
    aes(
      y = stat(count), 
      label = scales::percent(stat(prop), accuracy = 0.1), 
      group = key
    ), 
    stat = "count", 
    vjust = -0.5,
    size = 3
  ) +
  facet_wrap(~ key,) +
  scale_y_continuous(limits = c(0,400)) +
  theme(
    legend.position = "none",
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank()
  ) +
  labs(x = NULL, y = "Number of articles\n")

# Save
ggsave(
  filename = file.path(fig_output, "paper_fig-1a_indicator-prevalence.jpg"),
  plot = p1,
  width = 6, 
  height = 6
)
```


## Indicators over time

```{r}
# Table
tbl <- 
  articles %>% 
  filter(isEnglish & isResearch) %>%
  gather("indicator", "value", !!! indicator_vars_binary) %>% 
  group_by(Year, indicator) %>% 
  count(value) %>% 
  mutate(
    prop = n / sum(n),
    prop_formatted = scales::percent(prop),
    n_prop = sprintf("%.f (%s)", n, prop_formatted)
  ) %>% 
  filter(value) %>% 
  mutate(
    sd = sqrt(prop * (1 - prop) / n),
    ymin = max(0, prop - sd * qnorm(0.975)),
    ymax = min(1, prop + sd * qnorm(0.975)),
  ) %>% 
  ungroup() %>% 
  mutate(indicator = fct_relevel(indicator, "Data")) %>% 
  mutate(indicator = fct_relevel(indicator, "Protocol", after = 4)) %>% 
  mutate(indicator = recode(indicator, !!! indicator_vars_dic))

# Plot
p2 <- 
  tbl %>% 
  ggplot(aes(x = Year, y = prop, color = indicator, fill = indicator)) +
  geom_line(size = 1) +
  geom_ribbon(aes(ymin = ymin, ymax = ymax), alpha = 0.5, linetype = 2) +
  guides(fill = FALSE, color = FALSE) +
  facet_wrap(~ indicator) +
  scale_y_continuous(
    name = "Proportion with indicator (%)\n", 
    labels = scales::percent, 
    limits = c(0, 1)
  ) +
  coord_cartesian(expand = F) +
  theme(
    panel.grid.minor = element_blank()
  )

# Save
ggsave(
  filename = file.path(fig_output, "paper_fig-1b_indicator-over-time.jpg"),
  plot = p2,
  width = 8, 
  height = 6
)
```

## Join

```{r}
cowplot::plot_grid(
  p1, p2,
  labels = "auto",
  rel_widths = c(2, 3),
  scale = 0.9
)

ggsave(
  filename = file.path(fig_output, "paper_fig-1_indicators-across-levels.jpg"),
  width = 14, 
  height = 8
)
```


***


# Extended Fig 1: Sharing type

```{r}
# Table
tbl <- 
  articles %>% 
  filter(Reviewer != "JW" & isEnglish & isResearch) %>% 
  select(Sample_ID, FundingType, DataShareCategory) %>%
  drop_na() %>% 
  mutate(FundingType = recode(FundingType, !!! fund_dic)) %>% 
  gather(key, value, -Sample_ID) %>% 
  mutate(key = recode(key, !!! indicator_vars_dic)) %>% 
  filter(!value %in% c("No", "No statement")) %>% 
  group_by(key) %>% 
  add_count(N = n()) %>% 
  ungroup() %>% 
  separate_rows(value, sep = "; ") %>% 
  group_by(Sample_ID) %>% 
  distinct(value, .keep_all = T) %>% 
  ungroup() %>% 
  select(-Sample_ID) %>% 
  group_by(N, key) %>%
  count(value) %>% 
  mutate(p = scales::percent(n / N, accuracy = 0.1)) %>% 
  ungroup() %>% 
  arrange(n)

# Plot
tbl %>% 
  ggplot(aes(x = fct_inorder(value), y = n, fill = key, color = key)) +
  geom_segment(aes(xend = value, y = 0, yend = n), size = 1) +
  geom_point(size = 3) +
  geom_text(aes(y = n, label = p), hjust = -0.3, size = 3) +
  facet_wrap(~ key, scales = "free") +
  scale_y_continuous(expand = expansion(mult = c(0, .17))) +
  labs(x = NULL, y = "\nNumber of articles") +
  coord_flip() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_line(size = 0.1),
    panel.grid.minor.x = element_blank(),
    axis.text.y = element_text(margin = margin(r = 3)),
    axis.text.x = element_text(margin = margin(t = 3)),
    axis.ticks = element_blank(),
    axis.title = element_text(face = "bold"),
    legend.position = "none"
  )

# Save
ggsave(
  filename = file.path(fig_output, "paper_extended-fig-1_sharing-type.jpg"),
  width = 10, 
  height = 4
)
```


***


# Supplementary Table 1: PMC vs non-PMC

All.

```{r}
# Indicators for all articles
all <- 
  articles %>% 
  filter(Reviewer != "JW" & isEnglish) %>%
  mutate(is_PMC = is.na(PMCID)) %>% 
  gather(indicator, vals, COI, Funding) %>% 
  mutate(indicator = recode(indicator, !!! indicator_vars_dic)) %>% 
  group_by(indicator) %>% 
  mutate(p_val = fisher.test(vals, is_PMC)$p.value) %>% 
  group_by(is_PMC, indicator, p_val) %>% 
  count(vals) %>% 
  mutate(
    p = scales::percent(n / sum(n), accuracy = 0.1),
    p_text = sprintf("%.f (%s)", n, p)
  ) %>% 
  select(-n, -p) %>% 
  ungroup() %>% 
  mutate(is_PMC = if_else(is_PMC, "PMC", "non-PMC")) %>% 
  pivot_wider(names_from = is_PMC, values_from = p_text) %>% 
  mutate(vals = if_else(vals, "Yes", "No")) %>% 
  select(-p_val, p_val)

# Indicators for research articles
research <- 
  articles %>% 
  filter(Reviewer != "JW" & isEnglish & isResearch) %>%
  mutate(is_PMC = is.na(PMCID)) %>% 
  gather(indicator, vals, Data, Code, Protocol, Novel, Replication) %>% 
  mutate(indicator = fct_inorder(indicator)) %>% 
  mutate(indicator = recode(indicator, !!! indicator_vars_dic)) %>% 
  group_by(indicator) %>% 
  mutate(p_val = fisher.test(vals, is_PMC)$p.value) %>% 
  group_by(is_PMC, indicator, p_val) %>% 
  count(vals) %>% 
  mutate(
    p = scales::percent(n / sum(n), accuracy = 0.1),
    p_text = sprintf("%.f (%s)", n, p)
  ) %>% 
  select(-n, -p) %>% 
  ungroup() %>% 
  mutate(is_PMC = if_else(is_PMC, "PMC", "non-PMC")) %>% 
  pivot_wider(names_from = is_PMC, values_from = p_text) %>% 
  mutate(vals = if_else(vals, "Yes", "No")) %>% 
  select(-p_val, p_val)

# Join
indicator_by_pmc <- bind_rows(all, research)

# Save
write_csv(
  indicator_by_pmc, 
  "../../output/table_output/paper_indicator-by-pmc.csv"
)

# Print
indicator_by_pmc %>% 
  sable() %>% 
  collapse_rows(columns = c(1, 5), valign = "top")
```


***


# Documentation {.tabset}

## Session Info

```{r session_info, echo=FALSE}
print(sessionInfo(), locale = F)
```


## References

```{r refs, echo=FALSE}
(.packages()) %>% sort %>% lapply(citation) %>% lapply(c) %>% unique
```
